# Importing the original 2009 to 2019 YRBSS data
# Aaron Cromar
# 08-31-2023

---
title: "Importing YRBS Data"
author: "Aaron Cromar"
date: "2023-08-31"
output:
  pdf_document: default
  html_document: default
  word_document: default
---


# Loading R Packages
```{r setup, include = FALSE}
library(conflicted)
library(rmarkdown)
library(tidyverse)
library(haven)
library(foreign)
library(knitr)
library(usethis)
library(janitor)
```

# Importing .dat files into R for YRBS 2019
```{r}
raw2019YRBS <- read_sav("C:/Users/acrom/Google Drive/ASU Files/ASU WORK PRODUCT_CONHI Program/ASU-NUR CE Prep/YRBS Datasets/ADC Dissertation-YRBS Data Sets/2019YRBS-SPSS.sav")
View(raw2019YRBS)
```

# Print Head and Summary of 2019 YRBS Data Frame
```{r}
print("Top 6 Entries of data frame:")
head(raw2019YRBS)
print("summary:")
summary(raw2019YRBS)
```

# Importing .dat files into R for YRBS 2017
```{r}
raw2017YRBS <- read_sav("C:/Users/acrom/Google Drive/ASU Files/ASU WORK PRODUCT_CONHI Program/ASU-NUR CE Prep/YRBS Datasets/ADC Dissertation-YRBS Data Sets/2017YRBS-SPSS.sav")
View(raw2017YRBS)
```

# Print Head and Summary of 2017 YRBS Data Frame
```{r}
print("Top 6 Entries of data frame:")
head(raw2017YRBS)
print("summary:")
summary(raw2017YRBS)
```

# Importing .dat files into R for YRBS 2015
```{r}
raw2015YRBS <- read_sav("C:/Users/acrom/Google Drive/ASU Files/ASU WORK PRODUCT_CONHI Program/ASU-NUR CE Prep/YRBS Datasets/ADC Dissertation-YRBS Data Sets/2015YRBS-SPSS.sav")
View(raw2015YRBS)
```

# Print Head and Summary of 2015 YRBS Data Frame
```{r}
print("Top 6 Entries of data frame:")
head(raw2015YRBS)
print("summary:")
summary(raw2015YRBS)
```

# Importing .dat files into R for YRBS 2013
```{r}
raw2013YRBS <- read_sav("C:/Users/acrom/Google Drive/ASU Files/ASU WORK PRODUCT_CONHI Program/ASU-NUR CE Prep/YRBS Datasets/ADC Dissertation-YRBS Data Sets/2013YRBS-SPSS.sav")
View(raw2013YRBS)
```

# Print Head and Summary of 2013 YRBS Data Frame
```{r}
print("Top 6 Entries of data frame:")
head(raw2013YRBS)
print("summary:")
summary(raw2013YRBS)
```

# Importing .dat files into R for YRBS 2011
```{r}
raw2011YRBS <- read_sav("C:/Users/acrom/Google Drive/ASU Files/ASU WORK PRODUCT_CONHI Program/ASU-NUR CE Prep/YRBS Datasets/ADC Dissertation-YRBS Data Sets/2011YRBS-SPSS.sav")
View(raw2011YRBS)
```

# Print Head and Summary of 2011 YRBS Data Frame
```{r}
print("Top 6 Entries of data frame:")
head(raw2011YRBS)
print("summary:")
summary(raw2011YRBS)
```

# Importing .dat files into R for YRBS 2009
```{r}
raw2009YRBS <- read_sav("C:/Users/acrom/Google Drive/ASU Files/ASU WORK PRODUCT_CONHI Program/ASU-NUR CE Prep/YRBS Datasets/ADC Dissertation-YRBS Data Sets/2009YRBS-SPSS.sav")
View(raw2009YRBS)
```

# Print Head and Summary of 2009 YRBS Data Frame
```{r}
print("Top 6 Entries of data frame:")
head(raw2009YRBS)
print("summary:")
summary(raw2009YRBS)
```

# Summary of Raw 2009-2019 YRBS Data Waves
## Raw2019YRBS: 13677 x 235
## Raw2017YRBS: 14765 x 240
## Raw2015YRBS: 15624 x 239
## Raw2013YRBS: 13583 x 213
## Raw2011YRBS: 15425 x 207
## Raw2009YRBS: 16410 x 208